\bf Calendar-based graphics for visualizing people's daily schedules

### 2018-10-17 main.Rmd
rewrote letter value plot

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth. Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function. Section \ref{sec:variations} presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the advantages and disadvantages of the method.

# Construction 
\label{sec:algorithm}

@@

# Options
\label{sec:opt}

@@

## Layouts

@@

## Polar transformation

@@

## Scales

@@

## Orientation

@@

## Language support

@@

# Variations
\label{sec:variations}

## Overlaying and faceting subsets

@@

```{r facet, fig.height = 11, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
```

## Different types of plots

@@

@@

The case study uses individual-level energy usage from Australia. The dataset contains half-hourly electricity consumption data of 4 households over different time periods, with `r format(NROW(elec), big.mark = ",")` observations. The goal is to sketch typical profiles as well as unusual patterns in order to better understand each household's daily schedules.

Firstly, a letter value plot [@hofmann2017letter] shown as Figure \ref{fig:dow} helps grasp an overview of the variability between households across days of week. Household 2 has seen a great deal of variability in daily energy consumption, depicting relatively large variations on Thursdays. By contrast, household 4 consumes much less energy with constant variations. The dark blue line (representing median) highlights more energy consumed by households 1, 2, and 4 over the weekends relative to weekdays.

```{r dow, fig.height = 7, fig.cap = "Letter value plot of daily energy usage against day of week by households. The enery consumed by household 2 exhibits much larger variability, compared to the rest households. Households 1, 2, and 4 use mose energy over the weekends."}
```

Then a snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days. Figure \ref{fig:hod} gives a more detailed depiction of each household's life style. Household 1 appears an early bird, starting the day before 6 and going back home around 18 on workdays. There is an uncommon pattern occurring to household 2 only, with a cluster of points concentrated above 1 kWh across all time of day. It perhaps alerts household 2 to potential electricity leakage. Household 3 shows persistent trimodel behaviors across workdays and non-workdays. Nevertheless, the non-work life of household 4 distinguishes from the workdays when most energy is consumed after 18.

```{r hod, fig.height = 6, fig.cap = "Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household. A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day."}
```

Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households. Is there any untold story from these plots? Figure \ref{fig:household13} lays out the data points at the finest time resolution in a $3 \times 4$ monthly grid, for household 1 and 3. Household 1 was on vacation over three weeks of June, and household 2 was also away for holidays from September to early October and late June to mid July in the past year. This calendar plot evidently fills the missing piece of the picture.

```{r household13, fig.height = 11, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
```

@@

The focus is on the western calendar layout, because most countries have adopted this format. However the language of labels would differ across countries, and so support was added for changing the label language simply.

This sort of layout will be useful for studying consumer trends, or human behavior, such as pedestrian patterns or pollution peaks. It will not be so useful for physical patterns like temperature, which are not typically affected by human activity. The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data. Analysts would still be advised to plot overall summaries and deviations, in order to study general trends.

The layout is achieved by utilizing linear combinations of temporal components and measured variables. This provides the data with the most screen real estate. It could be useful to develop this into a fully-fledged faceting method, with formal labels and axes. This is a future goal.

>>>>

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth. Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function. Section \ref{sec:variations} presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

# Creating a calendar display
\label{sec:algorithm}

## Data transformation

@@

## Options
\label{sec:opt}

@@

### Layouts

@@

### Polar transformation

@@

### Scales

@@

### Orientation

@@

### Language support

@@

## Variations
\label{sec:variations}

### Overlaying and faceting subsets

@@

```{r facet, fig.height = 10, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
```

### Different types of plots

@@

## Interactivity

@@

The use of the calendar display is illustrated on smart meter energy usage from four households in Melbourne, Australia. Individuals can download their own data from the energy company, and these four households are the data of colleagues of the authors. The calendar display is useful to help people understand their energy use. The data contains half-hourly electricity consumption in 2017 and 2018. The analysis begins by looking at the distribution over days of week, then time of day split by workdays and non-work days, followed by the calendar display to inspect the daily schedules.

Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots [@hofmann2017letter]. As a variant of boxplots, letter value plots represent each pair of adjacent letter values using one box, instead of whiskers. Figure \ref{fig:dow} demonstrates three letter values and outliers marked as asterisks, with a line representing the median (thus M) and the innermost boxes corresponding to the fourth (F) and the eighth quantiles (E). Household 3, with a family size of one couple and two kids, uses more energy over the weekdays given the medians, compared to the others consisting of one or two people. Whereas, household 2 displays a great deal of variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends. The other two households (1 and 4) tend to consume more energy on average with increased variations during the weekends relative to the weekdays, as a reflection of work and leisure pattern.

```{r dow, fig.height = 7, fig.cap = "Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates. Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size. By constrast, household 2 sees considerally large variability."}
```

A snapshot is taken at a different time resolution to picture individual's energy usage against time of day, across workdays and non-work days. Figure \ref{fig:hod} gives a more detailed depiction of each household's life style. Household 1 appears an early bird, starting the day before 6 and going back home around 18 on workdays. There is an uncommon pattern occurring to household 2 only, with a cluster of points concentrated above 1 kWh across all time of day. It perhaps alerts household 2 to potential electricity leakage. Household 3 shows persistent trimodel behaviors across workdays and non-workdays. Nevertheless, the non-work life of household 4 distinguishes from the workdays when most energy is consumed after 18.

```{r hod, fig.height = 6, fig.cap = "Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household. A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day. (too generic, when they get up and different between work day and non-work days)"}
```

Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical toolkits, are useful for summarizing overall deviations across days and households. Is there any untold story from these plots? Figure \ref{fig:household13} lays out the data points at the finest time resolution in a $3 \times 4$ monthly grid, for household 1 and 3. Household 1 was on vacation over three weeks of June, and household 2 was also away for holidays from September to early October and late June to mid July in the past year. This calendar plot evidently fills the missing piece of the picture. Also weekday and weekend

```{r h12, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
```

```{r h34, fig.height = 10, fig.cap = "Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \\ref{fig:dow} and \\ref{fig:hod}."}
```

@@

The methodology creates the western calendar layout, because most countries have adopted this format. The main difference between countries is the use of different languages for labelling, which is supported by the software. Layouts beyond the western calendar could be achieved by the same modulars arithmetic approach.

The calendar layout will be useful for studying consumer trends and human behavior. It will not be so useful for physical patterns like climate, which are not typically affected by human activity. The layout does not replace traditional displays, but serves to complement to further tease out structure in temporal data. Analysts would still be advised to plot overall summaries and deviations, in order to study general trends.

The layout is a type of faceting and could be useful to develop this into a fully-fledged faceting method, with formal labels and axes. This is a future goal.

### 2018-10-17 main.tex
rewrote letter value plot

The remainder of the paper is organized as follows. Section
\ref{sec:algorithm} demonstrates the construction of the calendar layout
in depth. Section \ref{sec:opt} lists and describes the options that
come with the \texttt{frame\_calendar} function. Section
\ref{sec:variations} presents some variations of its usage. Graphical
analyses of sub-daily people's activities are illustrated with a case
study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses
the advantages and disadvantages of the method.

\hypertarget{construction}{%
\section{Construction}\label{construction}}

@@

@@

\hypertarget{options}{%
\section{Options}\label{options}}

@@

\hypertarget{layouts}{%
\subsection{Layouts}\label{layouts}}

@@

\hypertarget{polar-transformation}{%
\subsection{Polar transformation}\label{polar-transformation}}

@@

\hypertarget{scales}{%
\subsection{Scales}\label{scales}}

@@

\hypertarget{orientation}{%
\subsection{Orientation}\label{orientation}}

@@

\hypertarget{language-support}{%
\subsection{Language support}\label{language-support}}

@@

\hypertarget{variations}{%
\section{Variations}\label{variations}}

@@

\hypertarget{overlaying-and-faceting-subsets}{%
\subsection{Overlaying and faceting
subsets}\label{overlaying-and-faceting-subsets}}

@@

\hypertarget{different-types-of-plots}{%
\subsection{Different types of plots}\label{different-types-of-plots}}

@@

@@

The case study uses individual-level energy usage from Australia. The
dataset contains half-hourly electricity consumption data of 4
households over different time periods, with 64,128 observations. The
goal is to sketch typical profiles as well as unusual patterns in order
to better understand each household's daily schedules.

Firstly, a letter value plot \citep{hofmann2017letter} shown as Figure
\ref{fig:dow} helps grasp an overview of the variability between
households across days of week. Household 2 has seen a great deal of
variability in daily energy consumption, depicting relatively large
variations on Thursdays. By contrast, household 4 consumes much less
energy with constant variations. The dark blue line (representing
median) highlights more energy consumed by households 1, 2, and 4 over
the weekends relative to weekdays.

@@

\caption{Letter value plot of daily energy usage against day of week by households. The enery consumed by household 2 exhibits much larger variability, compared to the rest households. Households 1, 2, and 4 use mose energy over the weekends.}\label{fig:dow}
\end{figure}

Then a snapshot is taken at a different time resolution to picture
individual's energy usage against time of day, across workdays and
non-work days. Figure \ref{fig:hod} gives a more detailed depiction of
each household's life style. Household 1 appears an early bird, starting
the day before 6 and going back home around 18 on workdays. There is an
uncommon pattern occurring to household 2 only, with a cluster of points
concentrated above 1 kWh across all time of day. It perhaps alerts
household 2 to potential electricity leakage. Household 3 shows
persistent trimodel behaviors across workdays and non-workdays.
Nevertheless, the non-work life of household 4 distinguishes from the
workdays when most energy is consumed after 18.

@@

\caption{Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household. A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day.}\label{fig:hod}
\end{figure}

Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical
toolkits, are useful for summarizing overall deviations across days and
households. Is there any untold story from these plots? Figure
\ref{fig:household13} lays out the data points at the finest time
resolution in a \(3 \times 4\) monthly grid, for household 1 and 3.
Household 1 was on vacation over three weeks of June, and household 2
was also away for holidays from September to early October and late June
to mid July in the past year. This calendar plot evidently fills the
missing piece of the picture.

@@

{\centering \includegraphics[width=\textwidth]{figure/household13-1}

@@

\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:household13}
\end{figure}

>>>>

The remainder of the paper is organized as follows. Section
\ref{sec:algorithm} demonstrates the construction of the calendar layout
in depth. Section \ref{sec:opt} lists and describes the options that
come with the \texttt{frame\_calendar} function. Section
\ref{sec:variations} presents some variations of its usage. Graphical
analyses of sub-daily people's activities are illustrated with a case
study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses
the limitations of calendar displays and possible new directions.

\hypertarget{creating-a-calendar-display}{%
\section{Creating a calendar
display}\label{creating-a-calendar-display}}

@@

\hypertarget{data-transformation}{%
\subsection{Data transformation}\label{data-transformation}}

@@

\hypertarget{options}{%
\subsection{Options}\label{options}}

@@

\hypertarget{layouts}{%
\subsubsection{Layouts}\label{layouts}}

@@

\hypertarget{polar-transformation}{%
\subsubsection{Polar transformation}\label{polar-transformation}}

@@

\hypertarget{scales}{%
\subsubsection{Scales}\label{scales}}

@@

\hypertarget{orientation}{%
\subsubsection{Orientation}\label{orientation}}

@@

\hypertarget{language-support}{%
\subsubsection{Language support}\label{language-support}}

@@

\hypertarget{variations}{%
\subsection{Variations}\label{variations}}

@@

\hypertarget{overlaying-and-faceting-subsets}{%
\subsubsection{Overlaying and faceting
subsets}\label{overlaying-and-faceting-subsets}}

@@

\hypertarget{different-types-of-plots}{%
\subsubsection{Different types of
plots}\label{different-types-of-plots}}

@@

\hypertarget{interactivity}{%
\subsection{Interactivity}\label{interactivity}}

@@

The use of the calendar display is illustrated on smart meter energy
usage from four households in Melbourne, Australia. Individuals can
download their own data from the energy company, and these four
households are the data of colleagues of the authors. The calendar
display is useful to help people understand their energy use. The data
contains half-hourly electricity consumption in 2017 and 2018. The
analysis begins by looking at the distribution over days of week, then
time of day split by workdays and non-work days, followed by the
calendar display to inspect the daily schedules.

Figure \ref{fig:dow} shows the energy use across days of week in the
form of letter value plots \citep{hofmann2017letter}. As a variant of
boxplots, letter value plots represent each pair of adjacent letter
values using one box, instead of whiskers. Figure \ref{fig:dow}
demonstrates three letter values and outliers marked as asterisks, with
a line representing the median (thus M) and the innermost boxes
corresponding to the fourth (F) and the eighth quantiles (E). Household
3, with a family size of one couple and two kids, uses more energy over
the weekdays given the medians, compared to the others consisting of one
or two people. Whereas, household 2 displays a great deal of variability
in daily energy consumption with noticeable variations on Thursdays, and
much higher usage over the weekends. The other two households (1 and 4)
tend to consume more energy on average with increased variations during
the weekends relative to the weekdays, as a reflection of work and
leisure pattern.

@@

\caption{Letter value plots of daily energy usage against day of week for four households, with one line dislaying the median (M) and each box corresponding to the fourth (F) and eighth (E) paired quantile estimates. Suggested by the medians, household 3 uses more enery than the others on the weekdays, due to a large family size. By constrast, household 2 sees considerally large variability.}\label{fig:dow}
\end{figure}

A snapshot is taken at a different time resolution to picture
individual's energy usage against time of day, across workdays and
non-work days. Figure \ref{fig:hod} gives a more detailed depiction of
each household's life style. Household 1 appears an early bird, starting
the day before 6 and going back home around 18 on workdays. There is an
uncommon pattern occurring to household 2 only, with a cluster of points
concentrated above 1 kWh across all time of day. It perhaps alerts
household 2 to potential electricity leakage. Household 3 shows
persistent trimodel behaviors across workdays and non-workdays.
Nevertheless, the non-work life of household 4 distinguishes from the
workdays when most energy is consumed after 18.

@@

\caption{Scatterplot between time of day and half-hourly energy usage, constrasting workdays and non-work days for each household. A distinctive pattern exclusive to household 2 is popped out: a fair amount of points are clustered above 1 kWh over all time of day. (too generic, when they get up and different between work day and non-work days)}\label{fig:hod}
\end{figure}

Figure \ref{fig:dow} and \ref{fig:hod}, part of traditional graphical
toolkits, are useful for summarizing overall deviations across days and
households. Is there any untold story from these plots? Figure
\ref{fig:household13} lays out the data points at the finest time
resolution in a \(3 \times 4\) monthly grid, for household 1 and 3.
Household 1 was on vacation over three weeks of June, and household 2
was also away for holidays from September to early October and late June
to mid July in the past year. This calendar plot evidently fills the
missing piece of the picture. Also weekday and weekend

@@

{\centering \includegraphics[width=\textwidth]{figure/h12-1}

@@

\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h12}
\end{figure}

{\centering \includegraphics[width=\textwidth]{figure/h34-1}

\caption{Arranging sub-daily energy usage into monthly grids reveals the vacation time periods for households 1 and 3, which are hidden in Figure \ref{fig:dow} and \ref{fig:hod}.}\label{fig:h34}
\end{figure}
